---
title: "Final Memo"
author: "Margaret Reed"
date: "4/15/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Intro
Many of the ‘bluest’ cities and towns in America house a variety of universities 
and colleges. We began our data analysis investigation interested in 
investigating the impact Duke and other university students have on the 
elections of Durham and North Carolina. After parsing and exploring the data, we 
came away with a couple of questions to further pursue. How have NC Duke student 
registration numbers changed over time? How does this compare with the number of 
Dukes students who ended up voting in North Carolina? Likewise, we were also 
interested in looking into how political party affiliations and voting habits 
have changed over time. 


### Methodology
To begin, we needed data. We ended up using NC voter registration data and NC 
voter history data. However, nowhere in our data sets was there a column for 
Duke Student: Y/N. Therefore, we needed to find some sort of heuristic to 
approximate what constitutes a Duke student using the data we had available. 
After some trial and error, we ended up classifying Duke students as those whose 
mailing addresses were located on Duke's campus. The specific methodology is as 
follows:

- The mailing address must specify a voter lives in a dorm named: "Bassett", 
"Epworth", "Few Quad", "Gilbert Addoms", "Giles", "Jarvis", "Maxwell", 
"Mirecourt", "Pegram", "Randolph", "Randolph Hall", "Round Table", "Southgate", 
"Wayne Manor", "Wilson", or "Wannamaker". 

- Otherwise, the street name must be one of the following: "Duke University West 
Campus", "Duke University East Campus", "Towerview", "Alexander", or "Yearby".

- Finally, we also looked at two special cases: Duke-owned apartment 300 Swift, 
and the few outliers that were located on Campus Drive. 

All of these decisions were driven by an exploration of the data and looking at 
where high numbers of young people were concentrated near Duke's campus. Because 
our analysis only looks at registered voters in Durham county, NC- we are not 
considering any of the students who may have decided to vote in their home 
county or state. After classifying who the Duke students were (within our 
methodology), we were ready to begin our analysis. 


```{r library, include = F}
library(tidyverse)
library(janitor)
library(scales)
library(cowplot)
```

```{r load-data, include = F}
durham21 <- read_csv("data/duke_2021.csv")
durhamhist <- read_tsv("data/ncvhis32.txt")
```


```{r find-students, include = F}
duke_students <- durham21 %>%
  filter(
    street_name %in% c("DUKE UNIVERSITY WEST CAMPUS", 
                       "DUKE UNIVERSITY EAST CAMPUS", "TOWERVIEW", "ALEXANDER", 
                       "YEARBY") |
      (street_name %in% c("BASSETT", "EPWORTH", "FEW QUAD", "GILBERT ADDOMS", 
                          "GILES", "JARVIS", "MAXWELL", "MIRECOURT", "PEGRAM", 
                          "RANDOLPH", "RANDOLPH HALL", "ROUND TABLE", 
                          "SOUTHGATE", "WAYNE MANOR", "WILSON", "WANNAMAKER") && 
         street_type_cd == "DORM") |
      (street_name == "SWIFT" & house_num == "300") |
      (street_name == "CAMPUS"  & str_detect(house_num, "13\\d\\d"))
  )

full_data <- inner_join(
  duke_students, 
  durhamhist, 
  by = c("voter_reg_num" = "voter_reg_num")
  ) %>%
  mutate(
    registr_yr = lubridate::year(registr_dt),
    datev = as.Date(election_lbl, format = "%m/%d/%Y"),
    voted_year = lubridate::year(datev)
    )
```

``` {r count-data, include = F}
election_year <- full_data %>%
  filter(
    str_detect(election_desc, "GENERAL"),
    voted_year %in% c(2004, 2008, 2012, 2016, 2020)
    ) %>%
  select(voter_reg_num, voted_year) %>%
  distinct() %>%
  count(voted_year)
  

election_yr_reg <- duke_students %>%
  mutate(cancel_yr = lubridate::year(cancellation_dt)) %>%
  mutate(
    cancel_yr = ifelse(cancel_yr == 1900, 2023, cancel_yr)
  ) %>%
  transmute(voter_reg_num, year = map2(registr_yr, cancel_yr, seq, by = 1)) %>%
  unnest(cols = c(year)) %>%
  filter(year %in% c(2004, 2008, 2012, 2016, 2020)) %>%
  select(voter_reg_num, year) %>%
  distinct() %>%
  count(year)

newly_reg <- duke_students %>%
  filter(registr_yr %in% c(2004, 2008, 2012, 2016, 2020)) %>%
  select(voter_reg_num, registr_yr) %>%
  distinct() %>%
  count(registr_yr)
```

```{r prop-table, include = F}
prop_table <- election_year %>% 
  rename(year = voted_year, num_voted  = n) %>%
  full_join(
    election_yr_reg %>%
      rename(num_registered = n)
  ) %>%
  full_join(
    newly_reg %>%
      rename(year = registr_yr, num_actively_registered = n)
  ) %>%
  mutate(
    active_prop = num_actively_registered / num_registered,
    voted_prop = num_voted / num_registered
  )
```


### Analysis


We were interested in voter registration numbers over time. 

First, we looked into how many Duke students actively registered to vote during 
the presidential election years. Interestingly, it seems Duke students were the 
most motivated to register to vote in the 2012 election. It would be interesting 
to investigate further whether this is true- and dig into why, or if this can be 
attributed to the methodology we used to classify Duke students in 2012. 

Next, we looked at how many Duke students were registered to vote- whether that 
be because they actively registered or because a previous voter registration 
carried over. This value increased steadily over time, likely due to students 
forgetting to cancel their registration despite moving out of the county/state. 


```{r reg-active, echo = F}
p1 <- duke_students %>%
  ggplot(aes(x = registr_yr, y = n)) +
  geom_line(data = newly_reg, color = "green") +
  geom_point(data = newly_reg, color = "green", size = 3) +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 4)) +
  labs(
    x = NULL,
    y = NULL,
    title = "Number of Duke students who actively registered to vote in Durham, NC"
  ) +
  theme_minimal()
```

```{r reg-election, echo = F}
p2 <- duke_students %>%
  ggplot(aes(x = year, y = n)) +
  geom_line(data = election_yr_reg, color = "red") +
  geom_point(data = election_yr_reg, color = "red", size = 3) +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 4)) +
  labs(
    x = "Year",
    y = NULL,
    title = "Number of Duke students who were registered to vote in Durham, NC"
  ) +
  theme_minimal()
```

```{r plots-together, echo = F}
cowplot::plot_grid(p1, p2, ncol = 1)
```


We also looked at Duke students who voted in presidential election years. Voting 
numbers appeared to decrease from 2012 to 2020.


```{r voted-election, echo = F}
full_data %>%
  ggplot(aes(x = voted_year, y = n)) +
  geom_line(data = election_year, color = "blue") +
  geom_point(data = election_year, color = "blue", size = 3) +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 4)) +
  labs(
    x = "Year",
    y = "Number of students",
    title = "Number of Duke students who actually voted in Durham, NC"
  ) +
  theme_minimal()
```



```{r both-election, echo = F, warning = F}
ggplot() +
  geom_line(data = election_yr_reg, aes(x = year, y = n, color = "Registered")) +
  geom_line(data = election_year, aes(x = voted_year, y = n, color = "Voted")) +
  geom_line(data = newly_reg, aes(x = registr_yr, y = n, color = "Active Registration")) +
  geom_point(data = election_yr_reg, aes(x = year, y = n, color = "Registered")) +
  geom_point(data = election_year, aes(x = voted_year, y = n, color = "Voted")) +
  geom_point(data = newly_reg, aes(x = registr_yr, y = n, color = "Active Registration")) +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_x_continuous(breaks = seq(2012, 2020, by = 4), limits = c(2012, 2020)) +
  scale_color_manual(
    name = "Duke Students",
    breaks = c("Voted", "Registered", "Active Registration"),
    values = c("Voted" = "blue", "Registered" = "red", 
               "Active Registration" = "green")
  ) +
  labs(
    x = "Year",
    y = "Number of Duke students",
    title = "Number of Duke students who voted vs who registered to vote in Durham, NC"
  ) +
  theme_minimal()
```


We also graphed the proportion of registered Duke students who voted in Durham, 
NC. Likewise, the number of those who took advantage of their voter registration 
status and voted decreased from 2012 to 2020. 

```{r prop-plot, echo = F}
prop_table %>%
  drop_na(voted_prop) %>%
  ggplot(aes(x = as.factor(year), y = voted_prop)) +
  geom_col(fill = "blue4") +
  geom_text(
    aes(label = paste0(round(voted_prop, 2)*100, "%")), 
    vjust = 1.5, 
    color = "white"
    ) +
  scale_y_continuous(labels = label_percent()) +
  labs(
    x = "Year",
    y = "Percent of registered students",
    title = "Percent of registered Duke students who actually voted"
  ) +
  theme_minimal()
```

### Conclusion:

Based on our analysis from 2012 to 2020, Duke students are both registering to 
vote and voting less in presidential elections in Durham county, NC. Perhaps 
they are less motivated by the candidate or have less access to the polls. 
Alternatively, it is also possible that over time, more Duke students are 
choosing to vote in their home state/county rather than using their Duke 
residence. This might be because they feel they will have more of an impact in 
other states or because they don't want to change their registration. 


### Next Steps:

Regardless, with midterm elections coming up it is important for all the Duke 
students who are interested, to feel comfortable and empowered to register and 
then vote. Duke provides resources for students and faculty through the site 
[Duke Votes](https://vote.duke.edu/). According to their homepage, those who 
have questions can contact Duke Votes chair, Daisy Lane, at 
margarita.lane@duke.edu. They provide several 
[resources](https://vote.duke.edu/resources/) for those looking to get 
involved and/or register to vote. 

If we had more time, we would like to further hone our methodology for 
determining who is a Duke student and investigate what caused the surge in 
registrations and voting in 2012. 


### Further analysis

We also looked into our secondary question of how Duke students' political 
party affiliations changed over time.

```{r party-time-election-reg, echo = F}
duke_students %>%
  count(registr_yr, party_desc) %>%
  filter(registr_yr >= 2000,
         registr_yr %% 4 == 0) %>%
  ggplot(aes(x = registr_yr, y = n, color = party_desc)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 4)) +
  scale_color_manual(values = c("blue4", "green3", "red3", "grey")) +
  labs(
    x = "Year",
    y = "Number of Duke students registered",
    color = "Political party"
  ) +
  theme_minimal()
```

It seems that the number of Duke students registered as Democrats increased over 
time and the number of Duke students registered as unaffiliated decreased.

The same pattern can be seen in this filled bar chart.

```{r party-split-reg, echo = F}
duke_students %>%
  filter(registr_yr >= 2000, registr_yr %% 4 == 0) %>%
  ggplot(aes(x = registr_yr, fill = fct_rev(fct_infreq(party_desc)))) +
  geom_bar(position = "fill") +
  scale_x_continuous(breaks = seq(2000, 2020, by = 4)) +
  scale_fill_manual(values = c("green3", "red3", "blue4", "grey")) +
  labs(
    x = "Year",
    y = NULL,
    fill = "Party",
    title = "Duke students' registered party over time"
  ) +
  theme_minimal()
```

Additionally we looked at voting method over time:

```{r vote-method, echo = F}
full_data %>%
  filter(voted_year %in% seq(2012, 2020, 4)) %>%
  ggplot(
    aes(x = voted_year, fill = fct_rev(fct_infreq(voting_method)))
  ) +
  geom_bar(position = "fill") +
  scale_x_continuous(breaks = seq(2008, 2020, by = 4)) +
  colorblindr::scale_fill_OkabeIto() +
  labs(
    x = "Year", 
    y = NULL,
    fill = "Voting method",
    title = "Duke students' voting method over time"
  ) +
  theme_minimal()
```
